Shapes analysis for time series.

### 2024-05-22 NEURIPS24/appendix.tex
appendix add

\subsection{Settings for experiment: analysis of respiratory behavior in mice}
\label{appendix: mice_exp_setting}

\subsection{Settings for experiment: TS-LDDMM representation identifiability}
\label{appendix: settings_identifiability}
This experiment only involves the TS-LDDMM method in two different settings: 
\begin{itemize}
  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$. The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
\end{itemize}

\subsection{Settings for experiment: Robustness to irregular sampling}
This experiment is inspired from \cite{oh2024stable} where authors performs an extensive comparison of Neural Ordinary Differential Equations (Neural ODEs) methods~\cite{kidger2020neural}. Methods, network architectures, and training strategy are taken from its associated Github\footnote{\url{https://github.com/yongkyung-oh/Stable-Neural-SDEs}} and describe in what follows.
\label{appendix: settings_robustness}
\subsubsection{Benchmark methods}
In related work, we give an overview of Neurals ODEs methods and its relation with TS-LDDMM.  
\begin{itemize}
  \item RNN-based methods: Baseline reccurent neural networks including RNN~\cite{medsker1999recurrent}, LSTM~\cite{hochreiter1997long}, and GRU~\cite{chung2014empirical}.

  \item Attention-based methods: Multi-Time Attention Networks (MTAN)~\cite{shukla2021multi} and Multi-Integration Attention Module (MIAM)~\cite{lee2022multi}. Both handles multivariate time series irregularly sampled with attention mechanisms.
  \item Neural ODEs: ODE-LSTM~\cite{lechner2020learning} a form of Neural-ODEs used to learned continous latent representations. 
  \item Neural SDEs:  Neural SDE~\cite{liu2019neural} and Neural LNSDE~\cite{oh2024stable} have been proposed to model randomness in time-series using drift and diffusion terms as an extension of Neural-ODEs. 
  \item Shape-Analysis methods: TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}. From shape analysis, both methods learn representations by solving ODEs parametrized with Kernels. While both methods hamdle multivariate signal irregularly sampled, TS-LDDMM is specifically desgined for time series.
\end{itemize}

\subsection{Settings for experiment: Classification benchmark on regularly sampled datasets}
\label{appendix: settings_classification}

\subsection{Synthetic experiments}
\label{appendix:numerics_synthetic}
For any deformations generation in both experiments (well-specified and misspecified), we take $\sigma_{T,0},\sigma_{T,1},\sigma_x=(100,1,1)$ and $c_0,c_1=(1,0.1)$ for the kernel $K_\msg$
and $\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}=(2,1,2,0.6) $ for the varifold kernels $k_\pos,k_\dir$ related to the loss $\scrl$.

In both experiments, we have nb\_steps=300 abd $\eta_M=0.1$.

\subsection{Mouse experiments}
The number of steps is larger in the second experiment (before/after injection) because the deformations are sharper.

\subsection{Classification experiments}
\label{appendix:classification_implementation}

We defined a default parametrization for all classifiers.

For classifiers: \href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.deep_learning.CNNClassifier.html#aeon.classification.deep_learning.CNNClassifier}{CNN}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.networks.ResNetNetwork.html#aeon.networks.ResNetNetwork}{ResNet}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.feature_based.Catch22Classifier.html#aeon.classification.feature_based.Catch22Classifier}{Catch22}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.distance_based.KNeighborsTimeSeriesClassifier.html#aeon.classification.distance_based.KNeighborsTimeSeriesClassifier}{DTW-KNN}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.convolution_based.RocketClassifier.html#aeon.classification.convolution_based.RocketClassifier}{Rocket}
we used the aeon\footnote{https://www.aeon-toolkit.org/en/stable/index.html} implementations with their default settings.

For Tloss-SVC we used the implementation provided on github\footnote{https://github.com/mqwfrog/ULTS} 
with the following parameters for learning representations: batch\_size: 10, channels: 40, depth: 10, nb\_steps: 200, in\_channels: 1, 
kernel\_size: 3, lr: 0.001, nb\_random\_samples: 10, negative\_penalty: 1, out\_channels: 320, reduced\_size: 160. 
We used the Support Vector Classifier (\href{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html#sklearn.svm.SVC}{SVC})
from scikit-learn with thee regularization term C: 1. Others parameters are set to default.

For TS-LDDMMM-SVC, all kernels' parameters et optimizer parameter are presented in \Cref{appendix:table:set_up_exp_practical}.
As well, we used the Support Vector Classifier from scikit-learn with thee regularization term C: 1. Others parameters are set to default.

 \begin{table}[hbt!]
  \caption{Parameters used in all the experiments.
   For synthetic data, $K_\msg$ refers to the kernel used in the generation, which is the same for the estimation only in the well-specified case. $\bar{l}$ refers to the average time series length and $N_d$ refers to the number of dimensions.}
  \centering
  \begin{tabular}{lllll}

            \toprule 
            objects & Optimizer &$k_\pos,k_\dir$ & $K_\msg $ \\
            \midrule
            Parameter & (nb\_steps,$\eta_M$) &$(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$ & $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)$ \\
            \midrule
            Synthetic data well-specified &(400,0.1) &$(2,1,2,0.6)$& $(1,0.1,100,1,1)$   \\
            Synthetic data misspecified  & (400,0.1)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
            Mouse before injection  & (400,0.3)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
            Mouse before/after injection  & (400,0.3)  & $ (5,1,5,0.6)$ &$(1,0.1,150,1,1)$  \\
            Classification  & (400,0.1)  & $(2,N_d,2,N_d)$  & $(1,0.1,0.33\bar{l},1,N_d)$  \\

            \bottomrule
          \end{tabular}
        \label{appendix:table:set_up_exp_practical}
    \end{table}

@@

\section{Experiments}

\subsection{Identifiability experiments}
\label{appendix:identifiability}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
  \vspace{-2.5em}
  \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
  \label{fig:exemple_synthetic}
  \vspace{-1em}
\end{figure}

@@

% \begin{subfigure}{0.7\linewidth}
%   \centering
%   \resizebox{\columnwidth}{!}{%
%      \begin{tabular}{lrrrr}
%      \toprule
%      $\sigma_{T,0} \backslash \sigma_x$  & 0.1 & 1 & 100 \\
%      \midrule
%      1 & 2e+0 & 4e-2 & 4e-2 \\
%      10 & 3e-4 & 1e-4  & 2e-4 \\
%      50 & 1e-5 & 1e-5  & 1e-5 \\
%      100 & 4e-6 & 4e-6 & 4e-6 \\
%      200 & 7e-4 & 7e-4 & 7e-4 \\
%      300 & 4e-3 & 4e-3 & 4e-3 \\
%      \bottomrule
%      \end{tabular}
%      %
%      }
%  \caption{Values of $\scrl(y^*,\varphi^{\{\hat{v}_0\}}.\msx)$ where $y^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
% \end{subfigure}

@@

First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
Given $m_{s}\in \Nset_{>0}$ and $t_{a},s_{a}>0$, we randomly generate initial momentums $\mathbf{\alpha}^*=(\alpha_k^*)_{k\in[\mathbf{n}_0]}$ with the following sampling, called Gen($m_s,t_a,s_a$):
For any $k\in[\mathbf{n}_0]$, $\mathbf{\alpha}_k'$ is sampled according to a Gaussian normal distribution $\mathcal{N}(0_{d+1},I_{d+1})$.
Then, $(\alpha_k')_{k\in[\mathbf{n}_0]}$ is regularized by a rolling average of size $m_{s}$, we get $\bar{\mathbf{\alpha}}'=(\bar{\alpha}_k')_{k\in[\mathbf{n}_0]}$.
Finally, we normalize $\bar{\mathbf{\alpha}}'$ to derive $\mathbf{\alpha}^*$ such that $|([\alpha_k^*]_t)_{k\in[\mathbf{n}_0]}|=t_{\text{amp}}$ and $|([\alpha_k^*]_s)_{k\in[\mathbf{n}_0]}|=s_{\text{amp}}$ for any $k\in[\mathbf{n}_0]$, denoting by $[\alpha_k^*]_t,[\alpha_k^*]_s$ the time and space coordinates of $\alpha_k^*$ respectively.
Note that the regularizing step $(\mathbf{\alpha}_k')_{k\in[\mathbf{n}_0]}\to \bar{\mathbf{\alpha}}' $ is necessary to obtain realistic deformations which take into account the regularity induced by the RKHS $\msv$.
%  uniformly on the set $\mss_M=\{ (\alpha_k)_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}:\eqsp |\alpha_k|=M\}  $ with $M>0,\mathbf{n}_0\in \Nset_{>0}$,
%  then we regularize $\mathbf{\alpha}$ by convolving
%   with ... to get the generation parameter $\mathbf{\alpha}^*$.
Then, using $v_0(\mathbf{\alpha}^*,\msx)$ as defined in \eqref{eq:def_v0} with initial momentums $\mathbf{\alpha}^*$ and control points $\msx$, we apply the induced deformation $\varphi^{\{v_0\}} $ by \eqref{eq:integration} to $\msx$ and obtain $\varphi^{\{v_0\}}.\msx$.
Finally, we solve \eqref{eq:relaxation} to recover an estimation $\hat{\mathbf{\alpha}}$ of $\mathbf{\alpha}^*$ and report the average relative error (ARE) $|v_0(\hat{\mathbf{\alpha}},\msx)-v_0(\mathbf{\alpha}^*,\msx)|_\msv/|v_0(\mathbf{\alpha}^*,\msx)|_\msv$ on 50 repetitions.
This procedure is performed for any $m_{s},t_{a},s_{a}\in \{10,50,100\}\times \{5,10,15,20\}^2 $.
Mean, standard deviation, and maximum of the ARE on all these hyperparameters choices are respectively $\mathbf{0.10, 0.03, 0.17}$.
Therefore, the estimation procedure \eqref{eq:relaxation} offers a good approximation of the true parameter when the kernel $K_\msg$ is well specified.
We observe that the estimation is difficult when $t_a\ll s_a$ because the time series can be very noisy as illustrated in \Cref{fig:exemple_synthetic}: this impacts the Varifold loss which is sensitive to tangents.

@@

\subsection{Classification: Comparison with shape analysis methods}
\label{appendix:classification_shape_analysis}

@@

\section{Robustness to missing data}

%  [11:35, 31/01/2024] Thibaut Germain: avant/après :
%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,150,1,1)
%  Kl = TSGaussGaussKernel(5,1,5,0.6)
%  dataloss = VarifoldLoss(Kl)
%  schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
%  [11:35, 31/01/2024] Thibaut Germain: avant :
%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,100,1,1)
%  Kl = TSGaussGaussKernel(2,1,2,0.6)
%  dataloss = VarifoldLoss(Kl)
%  schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
% Let $s$ be a time series and $\msg(\tilde{s}) = (g_i)_{i \in [n]}$ be its discretized graph.
%  The graph $\msg(s)$ can be approximated as 
% the union of piecewise linear segments between time-consecutive samples, 
% and we associate to the approximated graph the oriented varifold:

% \begin{equation}
%   \mu_{\msg(\tilde{s})} : w \in \msw \mapsto \sum_{i=1}^{n-1} l_i\delta_{(c_i,\overrightarrow{v_i})}(w)
% \end{equation}
% where $\msw$ is a set of real valued test functions defined on $\Rset^{d+1} \times \mathbb{S}^d$, 
% $\delta_{(c_i,\overrightarrow{v_i})}(w) = w(c_i,\overrightarrow{v_i})$ is a Dirac delta function, 
% $c_i = (g_i + g_{i+1})/2$ (resp. $v_i = \| g_{i+1}-g_{i}\|$) is the center (resp. length) of the $i^{th}$ segment, 
% and $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$ is the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$.

% Assuming that the test functions' space $\msw$ is the RKHS associated with a $C^1$ positive definite kernel $k$ 
% that verifies \citep[Proposition 2 \& 4]{kaltenmark2017general}, the oriented varifold belongs to the dual space $\msw^*$. Additionally, 
% we can define a distance $d_{\msw^*}$ between signals' graph sample sets, which, thanks to the kernel reproducing property,
% has an explicit formulation:

% \section{Continuous normalizing flows}
% %Modeling Continuous Stochastic Processes with Dynamic Normalizing Flows time series generative
% CONTUNYIYS NORMALIZING FLOW bofbof
% Here, we present the framework of Continuous Normalizing Flows (CNF) and show how LDDMM can be seen as a particular case.
% \cite{salman2018deep} was already inspired from the work of shape analysis to present deep diffeomorphic normalizing flows.
% Discussion with Alain :
% Given a time-dependant vector field $v:[0,1]\times \Rset^{d'} \to \Rset^{d'} $, the integration of this vector field gives a flow $ \phi:[0,1]\times \Rset^{d'} \to \Rset^{d'}$, defined via the ordinary differential equation (ODE),
% \begin{equation}
%   \frac{\dd }{\dd \tau}\phi_\tau(x)=v(\tau,\phi_\tau(x)), \phi_0(x)=x \eqsp.
%  \end{equation}
% In \cite{chen2018neural}, they introduce Continous Normalizing flows for generative modeling.
%  By modeling $v=f_\theta$ is a neural networks and choosing a prior $X_0\sim\rho_0$ on the inital value, the law of $\phi_1(X_0)$ has the pullback density $\phi_1\# \rho_0$ and the parameter $\theta$ is optimized to maximize the log-likelihood of a given dataset $(y_i)_{i\in[N]}\in (\Rset^{d'})^N$ which is assumed to follow the law of $\phi_1(X_0)$.
% Instead of modeling the law of the end point $\phi_1(X_0) $, we can try to model the law of the whole path $(\phi_\tau(X_0))_{\tau\in[0,1]} $ having a density $p:[0,1]\times \Rset^d\to \Rset_{>0}$ such that $p_1=q, p_0=\rho_0$ where $q$ is the empirical distribution in the dataset, as proposed in \cite{lipman2022flow}.
% In so doing, the problem is ill-posed since $p_t$ can be whatever as soon as $0<t<1$, that is why we try to constraint the learned velocity field $v=f_\theta$ such that the path law $p$ is tractable.
% One way of doing it is to look for geodesics \eqref{eq:geodesics_original} as proposed in LDDMM, it corresponds to the case where $q=\sum_{i=1}^N \delta_{y_i}/N$ and $\rho_0=\sum_{i=1}^N \delta_{x_i}/N$ are empirical distribution and $y_i=\phi^v_1(x_i) $ for a given velocity field $v$.

% The goal is to target $\nu\in \mathcal{P}(\rmC^0([0,1]\times \Rset^d, \Rset^d)) $ by generating randomness on a initial value $x \sim \delta_x$
%  and by integrating a velocity field $v_\theta \in \msw \subset \rmC^0([0,1]\times \Rset^d, \Rset^d)$: 
%  \begin{equation}
%   \frac{\dd h}{\dd \tau}=v_\theta(\tau,h(\tau,x)), h(0,x)=x \eqsp,
%  \end{equation}
%  denoting by $\mu$ the law of $(h(\tau,x))_{\tau \in [0,1]}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \subsection{TTS}
% To circumvant this issue, we propose two methods :
%       \begin{enumerate}
%         \item \label{enum:first_method} [Time And Space (TAS)] Using a specific form of the RKHS's kernel in LDDMM, to stick to the classical shape analysis framework.
%         \item \label{enum:second_method} [Time Then Space (TTS)] Tackling the time axis and space axis separetely in the transformation.
%          This distangles time and space variability in the patterns, increasing the interpretability of the representation.
%           Although the related optimization is more complex than method \ref{enum:first_method}, the results is more robust to the choice of hyper-parameter related to the RKHS's kernel.
%       \end{enumerate}
%        In the following, we present both methods to highlights their pro and cons.

% As described in \Cref{section:optimization}, the method (TAS) learns $\Pi_{\gamma,f}$ directly, while (TTS) find the temporal reparametrisation $\Phi_f$ then the space transformation $\Psi_\gamma$.
%     % If we define a diffeomorpshim $\Phi_\gamma$ as $(\gamma,\Id_d) $, we have $\Phi_\gamma.G(s) =G(s).\gamma^{-1}$.
%     %  To increase the variability, we allow some deformations on the graph of the time series after being reparametrized in time, which allow us to recover every possible time seriess.
%     % Remark that, $G(\mathbf{s}_0\circ \psi^{-1})=\{(t,\mathbf{s}_0\circ \psi^{-1}(t)), t\in \msj \}= \{(\psi(t),\mathbf{s}_0(t)) ,\eqsp t\in \msi \}  $.
%     % [The action of the diffeomorpshim group 
%     % \begin{align}
%     %     &\mcd_*=\mcd(\Rset,\Rset)\times \{\phi\in \mcd(\Rset^{d+1},\Rset^{d+1}): \\
%     %     &\phi(t,x)=(t,f(t,x)),
%     %      (t,x)\in \Rset \times \Rset^{d}, f \in C^0(\Rset^{d+1}, \Rset^d) \} 
%     % \end{align}
%     % on the space of time series graph $\{G^{(a,b)}(s), s\in C^0((a,b), \Rset), (a,b)\in \Rset_+^2, a<b  \}$
%     % is homogene.]
%      %In the next section we define a metric and a norm on this space in order to find a representation $(\phi,\psi)$ with a minimal norm, and we show in (...) how to recover a proxy.
%      \paragraph{The representation as the solution of an optimization}
%      By \Cref{theorem:representation}, for any $\mathbf{s}_0\in C^0(\msi,\Rset^d)$, we can find $ F=(\gamma_j,f_j)_{j\in[N]}$ such that $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$ for any $j\in[N]$.
%      However, to get a meaningfull representation and to tackle the unicity problem, we should optimize the choice of $\mathbf{s}_0 $ and $F$, formally speaking, denoting by $R(\cdot) $ a regularization norm on $\mcd(\Rset^{d+1})$ and $L$ a loss on sets which will be specify later, we aim to solve
%      \begin{equation}
%       \label{eq:minimization}
%       \underset{\mathbf{s_0},(\gamma_j,f_j)_{j\in[N]}}{ \argmin} \sum_{j=1}^N \lambda R(\Pi_{\gamma_j,f_j})+ L(\msg(s^j),\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)) \eqsp ,
%     \end{equation} 
%     where $\lambda>0$ is a penalization factor related to $L$ to relax the condition $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$.

%     In practice, as already proposed in shape analysis, we perform an alternative minimization between $\mathbf{s_0}$ and $(\gamma_j,f_j)_j$.
%     The optimization on $(\gamma_j,f_j)_j$ is parrallelisable, for each couple $(\mathbf{s}_0,s^j)$, we should learn diffemorphisms of minimal norm satisfying a condition. 
%     To this end, we adapt the geodesic shooting method using LDDMM presented in \cite{durrleman2013sparse} to our special case.
%       Before to describe how to optimize the whole in \Cref{section:optimization}, we introduce LDDMM in the next section to specify the norm on diffemorphisms.

% \paragraph{Kernel choices}
% As depicted on figure (mettre ref), we can not use any kernel $K$ to apply the previous methodology to learn deformations on time series' graphs.

% In the following, we expose two choice of RKHS's kernel which answers this issue : (TAS) and (TTS).
%  We rely on the Gaussian kernel, denoting by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
% To apply (TAS) in LDDMM with $d'=d+1$, we use the following anisotropic Gaussian kernel  :
% \begin{align}
%   \label{eq:kernel_TAS}
%   &K_{\text{TAS}}((t,x),(t',x'))=\begin{pmatrix}
%     c_0K & 0 \\
%     0 & c_1 D 
%     \end{pmatrix} \eqsp , \\
%     &D=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
% \end{align}
% parametrized by the widths $\sigma_{T,0},\sigma_{T,1},\sigma_x>0$ and the constants $c_0,c_1>0$.
%  If we use this kernel, by considering \eqref{eq:specific_form}, we remark that the first coordinate of the velocity field $v_\tau$ only depends on the time variable $t$.
%   It implies that the final deformation will have the form $\phi^v=(\gamma,f) $ with  $f\in C^1(\Rset^{d+1},\Rset^d) $, and $\gamma\in \mcd(\Rset)$
%   since the first coordinate follows a LDDMM-ODE \eqref{eq:LDDMM_dynamic}.

% To apply (TTS), we use LDDMM on the time axis with $d'=1$ by taking $K_{\text{TTS}}^{\text{time}}=K_{\sigma_{T,0}}^{(1)}$, then on the space axis with $d'=d$ by choosing
% $K_{\text{TTS}}^{\text{space}}=K_{\sigma_{T,1}}^{(1)}\times K_{\sigma_x}^{(d)} $ where the widths are the same than in \eqref{eq:kernel_TAS}. The alternate optimization will be more detailed in \Cref{section:optimization}.

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

>>>>

@@

\section{Appendix for experiment: Analysis of respiratory behavior in mice}
\subsection{Settings}
\label{appendix: mice_exp_setting}

\section{Appendix for experiment: TS-LDDMM representation identifiability}
\subsection{Settings}
\label{appendix: settings_identifiability}
This experiment only involves the TS-LDDMM method in two different settings: 
\begin{itemize}
  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$. The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
\end{itemize}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
  \vspace{-1em}
  \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
  \label{fig:exemple_synthetic}
  \vspace{-1em}
\end{figure}

@@

@@

First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
Given $m_{s}\in \Nset_{>0}$ and $t_{a},s_{a}>0$, we randomly generate initial momentums $\mathbf{\alpha}^*=(\alpha_k^*)_{k\in[\mathbf{n}_0]}$ with the following sampling, called Gen($m_s,t_a,s_a$):
For any $k\in[\mathbf{n}_0]$, $\mathbf{\alpha}_k'$ is sampled according to a Gaussian normal distribution $\mathcal{N}(0_{d+1},I_{d+1})$.
Then, $(\alpha_k')_{k\in[\mathbf{n}_0]}$ is regularized by a rolling average of size $m_{s}$, we get $\bar{\mathbf{\alpha}}'=(\bar{\alpha}_k')_{k\in[\mathbf{n}_0]}$.
Finally, we normalize $\bar{\mathbf{\alpha}}'$ to derive $\mathbf{\alpha}^*$ such that $|([\alpha_k^*]_t)_{k\in[\mathbf{n}_0]}|=t_{\text{amp}}$ and $|([\alpha_k^*]_s)_{k\in[\mathbf{n}_0]}|=s_{\text{amp}}$ for any $k\in[\mathbf{n}_0]$, denoting by $[\alpha_k^*]_t,[\alpha_k^*]_s$ the time and space coordinates of $\alpha_k^*$ respectively.
Note that the regularizing step $(\mathbf{\alpha}_k')_{k\in[\mathbf{n}_0]}\to \bar{\mathbf{\alpha}}' $ is necessary to obtain realistic deformations which take into account the regularity induced by the RKHS $\msv$.

Then, using $v_0(\mathbf{\alpha}^*,\msx)$ as defined in \eqref{eq:def_v0} with initial momentums $\mathbf{\alpha}^*$ and control points $\msx$, we apply the induced deformation $\varphi^{\{v_0\}} $ by \eqref{eq:integration} to $\msx$ and obtain $\varphi^{\{v_0\}}.\msx$.
Finally, we solve \eqref{eq:relaxation} to recover an estimation $\hat{\mathbf{\alpha}}$ of $\mathbf{\alpha}^*$ and report the average relative error (ARE) $|v_0(\hat{\mathbf{\alpha}},\msx)-v_0(\mathbf{\alpha}^*,\msx)|_\msv/|v_0(\mathbf{\alpha}^*,\msx)|_\msv$ on 50 repetitions.
This procedure is performed for any $m_{s},t_{a},s_{a}\in \{10,50,100\}\times \{5,10,15,20\}^2 $.
Mean, standard deviation, and maximum of the ARE on all these hyperparameters choices are respectively $\mathbf{0.10, 0.03, 0.17}$.
Therefore, the estimation procedure \eqref{eq:relaxation} offers a good approximation of the true parameter when the kernel $K_\msg$ is well specified.
We observe that the estimation is difficult when $t_a\ll s_a$ because the time series can be very noisy as illustrated in \Cref{fig:exemple_synthetic}: this impacts the Varifold loss which is sensitive to tangents.

@@

\section{Appendix for experiment: Robustness to irregular sampling}
\label{appendix: robustness}

This experiment is inspired by \cite{oh2024stable} where the authors perform an extensive comparison of Neural Ordinary Differential Equations (Neural ODEs) methods~\cite{kidger2020neural}. We assess the classification
performances of several methods under regular sampling (0\% missing rate) and three irregular sampling regimes on 15 shape-based datasets (7 univariate \& 8 multivariate). Methods and training strategy are taken from its associated Github\footnote{\url{https://github.com/yongkyung-oh/Stable-Neural-SDEs}} and described in what follows. We conclude with the results, which show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%, 30\%, and 50\%.

\subsection{Benchmark methods}
\label{appendix: benchmark robustness}
In related work, we give an overview of Neurals ODEs methods and their relation with TS-LDDMM.  
\begin{itemize}
  \item RNN-based methods: Baseline reccurent neural networks including RNN~\cite{medsker1999recurrent}, LSTM~\cite{hochreiter1997long}, and GRU~\cite{chung2014empirical}.

  \item Attention-based methods: Multi-Time Attention Networks (MTAN)~\cite{shukla2021multi} and Multi-Integration Attention Module (MIAM)~\cite{lee2022multi}. Both handle multivariate time series irregularly sampled with attention mechanisms.
  \item Neural ODEs: ODE-LSTM~\cite{lechner2020learning} a form of Neural-ODEs used to learn continuous latent representations. 
  \item Neural SDEs:  Neural SDE~\cite{liu2019neural} and Neural LNSDE~\cite{oh2024stable} have been proposed to model randomness in time-series using drift and diffusion terms as an extension of Neural-ODEs. 
  \item Shape-Analysis methods: TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}. From shape analysis, both methods learn representations by solving ODEs parametrized with Kernels. While both methods handle multivariate signals irregularly sampled, TS-LDDMM is specifically designed for time series.
\end{itemize}

\subsection{Model architecture}

\paragraph{Neural ODEs methods}
As depicted in~\cite{oh2024stable}, any Neural ODEs layer in \Cref{appendix: benchmark robustness} is followed by an MLP with two fully connected layers with \texttt{ReLU} activations. The risk of overfitting and the model regularization are handled with a dropout rate of 10\% and an early-stopping mechanism, ceasing the training when the validation loss does not improve for 10 successive epochs.

\paragraph{TS-LDDMM and LDDMM}

\subsection{Protocol}

\subsection{Results}

\subsection{Appendix for experiment: Classification benchmark on regularly sampled datasets}
\subsection{Settings}
\label{appendix: settings_classification}

@@
